---
title: "OverReact Demo"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

This notebook demonstrates the main workflow in `OverReact`: cleaning inputs, building descriptive tables, estimating prevalence, fitting regression models, drawing forest plots, applying the package themes, and exporting outputs.

## Setup

```{r}
knitr::opts_chunk$set(fig.width = 8, fig.height = 4.8)

library(dplyr)
library(ggplot2)

repo_root <- normalizePath(
  if (basename(getwd()) == "demo") ".." else ".",
  winslash = "/",
  mustWork = TRUE
)

if (requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(repo_root, quiet = TRUE)
} else {
  r_files <- list.files(file.path(repo_root, "R"), pattern = "\\.R$", full.names = TRUE)
  invisible(lapply(r_files, source))
}

set.seed(123)
```

## Simulated Analysis Dataset

```{r}
n <- 1000

dat <- tibble::tibble(
  age_years = round(rnorm(n, mean = 50, sd = 15)),
  sex = factor(sample(c("Female", "Male"), n, replace = TRUE)),
  region = factor(sample(c("North", "Midlands", "South"), n, replace = TRUE, prob = c(0.30, 0.30, 0.40))),
  ethnicity = factor(sample(c("White", "Asian", "Black", "Mixed"), n, replace = TRUE, prob = c(0.70, 0.12, 0.10, 0.08))),
  deprivation = factor(sample(paste0("Q", 1:5), n, replace = TRUE), levels = paste0("Q", 1:5)),
  household_size = sample(1:6, n, replace = TRUE, prob = c(0.18, 0.28, 0.20, 0.17, 0.10, 0.07)),
  exercise_days = pmax(0, pmin(7, round(rnorm(n, mean = 3.5, sd = 1.8)))),
  occupation_group = factor(sample(c("Office", "Service", "Manual"), n, replace = TRUE, prob = c(0.45, 0.30, 0.25))),
  smoking_status = factor(sample(c("Never", "Former", "Current"), n, replace = TRUE, prob = c(0.55, 0.25, 0.20))),
  exposure_primary = rnorm(n),
  bmi = rnorm(n, mean = 27, sd = 4),
  survey_weight = runif(n, min = 0.7, max = 1.5),
  sample_date_chr = format(as.Date("2022-01-01") + sample(0:120, n, replace = TRUE), "%d/%m/%Y"),
  binary_1_2 = sample(c(1, 2), n, replace = TRUE, prob = c(0.25, 0.75)),
  binary_1_0 = sample(c(0, 1), n, replace = TRUE, prob = c(0.70, 0.30)),
  continuous_dirty = ifelse(runif(n) < 0.03, -1, rnorm(n, mean = 10, sd = 3))
)

linpred <- -1.1 +
  0.55 * dat$exposure_primary +
  0.03 * (dat$age_years - 50) +
  0.35 * (dat$sex == "Male") +
  0.45 * (dat$smoking_status == "Current") +
  0.25 * (dat$region == "North")

dat$outcome_bin <- rbinom(n, size = 1, prob = plogis(linpred))
dat$outcome_group <- factor(dat$outcome_bin, levels = c(0, 1), labels = c("Negative", "Positive"))
dat$outcome_cont <- 2 + 0.8 * dat$exposure_primary - 0.15 * (dat$sex == "Male") + rnorm(n)
dat$res <- as.character(dat$outcome_bin)
dat$predicted_flag <- as.integer(dat$exposure_primary + rnorm(n, sd = 0.8) > 0.6)

dat$region[sample(seq_len(n), 15)] <- NA
dat$smoking_status[sample(seq_len(n), 20)] <- NA

attr(dat$exposure_primary, "label") <- "Primary exposure score"
attr(dat$smoking_status, "label") <- "Smoking status"

cov_name_list <- list(
  sex = "Sex",
  smoking_status = "Smoking status",
  region = "Region",
  age_years = "Age (years)",
  bmi = "Body mass index",
  ethnicity = "Ethnicity",
  deprivation = "Deprivation quintile",
  household_size = "Household size",
  exercise_days = "Exercise days per week",
  occupation_group = "Occupation group"
)

adjusters <- c(
  "age_years",
  "sex",
  "region",
  "ethnicity",
  "deprivation",
  "household_size",
  "exercise_days",
  "occupation_group"
)

dat |>
  dplyr::select(outcome_group, age_years, sex, region, smoking_status, exposure_primary, bmi) |>
  head(8) |>
  knitr::kable()
```

## Cleaning Helpers

`reactCleaner()`, `binaryCleaner()`, `binaryCleaner_1_0()`, `continuousCleaner()`, `dateCleaner()`, and `checkConvertNumeric()` are intended for quick preprocessing before tabulation or modelling.

```{r}
cleaning_preview <- tibble::tibble(
  raw_binary_1_2 = dat$binary_1_2[1:8],
  binaryCleaner = binaryCleaner(dat$binary_1_2[1:8]),
  raw_binary_1_0 = dat$binary_1_0[1:8],
  binaryCleaner_1_0 = binaryCleaner_1_0(dat$binary_1_0[1:8]),
  raw_continuous = round(dat$continuous_dirty[1:8], 2),
  continuousCleaner = continuousCleaner(dat$continuous_dirty[1:8], floornumber = 0),
  reactCleaner_binary = reactCleaner(dat$binary_1_0[1:8], datatype = "binary", negval = 0, posval = 1)
)

cleaning_preview |>
  knitr::kable()
```

```{r}
date_preview <- tibble::tibble(
  raw_date = dat$sample_date_chr[1:8],
  dateCleaner = as.character(dateCleaner(dat$sample_date_chr[1:8], dateFormat = "%d/%m/%Y")),
  reactCleaner_date = reactCleaner(
    dat$sample_date_chr[1:8],
    datatype = "date",
    dateFormat = "%d/%m/%Y"
  )
)

date_preview |>
  knitr::kable()
```

```{r}
numeric_conversion_demo <- tibble::tibble(
  char_numeric = c("1", "2", NA, "4"),
  char_text = c("low", "medium", "high", "missing")
)

checkConvertNumeric(numeric_conversion_demo) |>
  knitr::kable()
```

## Cross Tabs and Table One

```{r}
smoking_by_sex <- crossTab(
  dat = dat,
  rowvar = "smoking_status",
  colvar = "sex",
  confint = FALSE,
  include_percentages = TRUE
)

smoking_by_sex |>
  knitr::kable()
```

```{r}
bmi_by_sex <- crossTabContinuous(
  dat = dat,
  rowvar = "bmi",
  colvar = "sex",
  summary_stat = "mean"
)

bmi_by_sex |>
  knitr::kable()
```

```{r}
multi_xtab <- crossTabMulti(
  dat = dat,
  rowvars = c("smoking_status", "bmi"),
  colvar = "sex",
  summary_stat = "mean",
  include_percentages = TRUE
)

multi_xtab |>
  knitr::kable()
```

```{r}
table_one <- tableOne(
  dat = dat,
  rowvars = c("age_years", "bmi", "smoking_status", "region"),
  colvar = "outcome_group",
  cov_names = cov_name_list,
  summary_stat = "mean",
  include_percentages = TRUE,
  statistical_test = TRUE
)

table_one |>
  head(10) |>
  knitr::kable()
```

## Prevalence Tables

```{r}
prevalence_tables <- makeTablesNew(
  dat = dat,
  result_var = "res",
  covariates = c("sex", "smoking_status", "region"),
  cov_name_list = cov_name_list,
  output_list = FALSE,
  percent = TRUE
)

prevalence_tables |>
  knitr::kable()
```

```{r}
stratified_prevalence <- stratifiedTables(
  dat = dat,
  result_var = "res",
  covariates = c("smoking_status"),
  cov_name_list = cov_name_list,
  strat_var = "sex",
  include_counts = TRUE
)

stratified_prevalence |>
  knitr::kable()
```

## Formatting and Epidemiology Utilities

```{r}
formatting_demo <- tibble::tibble(
  raw_value = c(pi, exp(1), 0.00042),
  specifyDecimal = specifyDecimal(c(pi, exp(1), 0.00042), k = 3),
  pValEpiConverter = pValEpiConverter(c(0.20, 0.02, 0.0001)),
  pvalAsterisker = pvalAsterisker(c(0.20, 0.02, 0.0001), return_p = TRUE)
)

formatting_demo |>
  knitr::kable()
```

```{r}
calculateEpiStats(
  prediction = dat$predicted_flag,
  outcome = dat$outcome_bin,
  returnAllStats = TRUE
)
```

## Individual Models and Effect Tables

```{r}
fit_glm <- glm(
  outcome_bin ~ exposure_primary + smoking_status + age_years + sex,
  data = dat,
  family = binomial()
)

fit_lm <- lm(
  outcome_cont ~ exposure_primary + smoking_status + age_years + sex,
  data = dat
)

or_table <- makeORTable(fit_glm)
rd_table <- makeRDTable(
  mod = fit_glm,
  variable_name = "smoking_status",
  ref_level = levels(dat$smoking_status)[1],
  dp = 3,
  data = dat,
  n_sim = 20
)

or_table |>
  knitr::kable()
```

```{r}
rd_table |>
  knitr::kable()
```

`plotReactForest()` accepts raw models, model lists, tidy plotting data frames, and `ModelMaker*` outputs.

```{r}
plotReactForest(
  fit_glm,
  variables = c("exposure_primary", "smoking_status"),
  title = "Forest plot from a logistic model"
)
```

```{r}
plotReactForest(
  fit_lm,
  variables = c("exposure_primary", "smoking_status"),
  title = "Forest plot from a linear model"
)
```

```{r}
rd_plot_data <- rd_table |>
  dplyr::mutate(
    Variable = "Smoking status",
    Category = Level
  )

plotReactForest(
  rd_plot_data,
  effect = "RD",
  variables = "Smoking status",
  title = "Risk-difference forest plot"
)
```

## Bulk Odds-Ratio Models

`ModelMakerMulti()` is the main batch OR/Beta workflow that does not rely on the RD-specific `data.table` path.

```{r}
multi_or <- suppressMessages(
  ModelMakerMulti(
    dat = dat,
    list_of_variables_of_interest = c("exposure_primary", "smoking_status", "bmi"),
    outcome = "outcome_bin",
    joint_adjustment_vars = c("age_years", "sex", "region"),
    ncores = 1
  )
)

multi_or$df_output |>
  head(10) |>
  knitr::kable()
```

```{r}
plotReactForest(
  multi_or,
  variables = c("Primary exposure score", "Smoking status"),
  adjustment_numbers = c("1", "2", "3", "4"),
  adjustment_descriptions = c(
    "Crude",
    "Level 1: + age",
    "Level 2: + age + sex",
    "Level 3: + age + sex + region"
  ),
  title = "Forest plot from ModelMakerMulti()"
)
```

## Forest Plots with 1 to 8 Adjustment Levels

The next helper builds model lists with exactly 1 through 8 plotted adjustment levels. This is the simplest way to demonstrate the full `plotReactForest()` API over a controlled range of adjustment counts.

```{r}
level_labels <- c(
  "Level 1: + age",
  "Level 2: + age + sex",
  "Level 3: + age + sex + region",
  "Level 4: + age + sex + region + ethnicity",
  "Level 5: + age + sex + region + ethnicity + deprivation",
  "Level 6: + age + sex + region + ethnicity + deprivation + household size",
  "Level 7: + age + sex + region + ethnicity + deprivation + household size + exercise days",
  "Level 8: full model"
)

make_adjustment_model_list <- function(k) {
  model_list <- vector("list", length = k)

  for (i in seq_len(k)) {
    rhs <- c("exposure_primary", "smoking_status", adjusters[seq_len(i)])
    model_list[[i]] <- glm(
      reformulate(rhs, response = "outcome_bin"),
      data = dat,
      family = binomial()
    )
  }

  names(model_list) <- paste0("adj_", seq_len(k))
  model_list
}

forest_plots_1_to_8 <- lapply(1:8, function(k) {
  plotReactForest(
    make_adjustment_model_list(k),
    variables = c("exposure_primary", "smoking_status"),
    adjustment_numbers = as.character(0:(k - 1)),
    adjustment_descriptions = level_labels[seq_len(k)],
    title = paste("Forest plot with", k, if (k == 1) "adjustment level" else "adjustment levels")
  )
})

for (p in forest_plots_1_to_8) {
  print(p)
}
```

## Themes, Palettes, and Scales

```{r}
palette_values <- tibble::tibble(
  palette_index = seq_len(5),
  hex = imperial_pal_discrete("default")(5)
)

palette_values |>
  knitr::kable()
```

```{r}
ggplot(dat, aes(x = sex, y = exposure_primary, fill = sex)) +
  geom_boxplot(width = 0.65) +
  scale_fill_new(palette = "cool") +
  theme_react() +
  labs(
    title = "theme_react() with scale_fill_new()",
    x = NULL,
    y = "Primary exposure score"
  )
```

```{r}
ggplot(dat, aes(x = age_years, y = outcome_cont, color = smoking_status)) +
  geom_point(alpha = 0.45) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.7) +
  scale_color_new(palette = "warm") +
  theme_mw() +
  labs(
    title = "theme_mw() with scale_color_new()",
    x = "Age (years)",
    y = "Continuous outcome"
  )
```

## Export Helpers

The save helpers below write to `tempdir()` so the demo remains self-contained and does not modify the repository.

```{r}
demo_output_dir <- file.path(tempdir(), "overreact-demo")
dir.create(demo_output_dir, recursive = TRUE, showWarnings = FALSE)

prevalence_plot <- ggplot(prevalence_tables, aes(x = Category, y = Prevalence, fill = Variable)) +
  geom_col(position = position_dodge(width = 0.8)) +
  geom_errorbar(
    aes(ymin = Lower, ymax = Upper),
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  scale_fill_new(palette = "default") +
  theme_react() +
  labs(
    title = "Prevalence estimates by category",
    x = NULL,
    y = "Prevalence (%)"
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

prevalence_plot
```

```{r}
saveREACTplot(
  p = prevalence_plot,
  figpath = demo_output_dir,
  filename = "prevalence_plot",
  width = 7,
  height = 4.5,
  savePDF = FALSE
)

saveREACTtable(
  tab = prevalence_tables,
  outpath = paste0(demo_output_dir, "/"),
  filename = "prevalence_table",
  save_rds = TRUE
)

saveREACT(
  file = prevalence_tables,
  figpath = demo_output_dir,
  outpath = paste0(demo_output_dir, "/"),
  filename = "prevalence_table_generic"
)

savePrettyExcelWorkbook(
  listOfTables = list(
    table_one = table_one,
    prevalence = prevalence_tables
  ),
  workbookName = "overreact_demo_tables",
  outpath = paste0(demo_output_dir, "/")
)

sort(list.files(demo_output_dir))
```

## Optional: RD-Aware Bulk Models and Automated Reports

`ModelMakerMultiRD()` and `run_auto_tableone_regressions()` are included below as reference examples. They are left unevaluated here to keep the demo fast, but the code shows the intended usage.

```{r}
#| eval: false
multi_rd <- ModelMakerMultiRD(
  dat = dat,
  list_of_variables_of_interest = c("exposure_primary", "smoking_status", "bmi"),
  outcome = "outcome_bin",
  joint_adjustment_vars = adjusters,
  incremental = TRUE,
  include_crude = TRUE,
  include_rd = TRUE,
  n_sim = 100,
  ncores = 1
)

plotReactForest(
  multi_rd,
  variables = c("exposure_primary", "smoking_status"),
  adjustment_numbers = c("0", as.character(seq_along(level_labels))),
  adjustment_descriptions = c("Crude", level_labels),
  title = "Forest plot from ModelMakerMultiRD()"
)

auto_results <- run_auto_tableone_regressions(
  dat = dat,
  outcome = "outcome_bin",
  covariates = adjusters,
  independent_variables = c("exposure_primary", "smoking_status", "bmi"),
  cov_names = cov_name_list,
  report_path = file.path(demo_output_dir, "auto_summary.md"),
  n_sim = 100
)

names(auto_results)
```
